---
title: "Week 5: Lab - Word Cloud Chapter Challenge"
author: "Ammar Ali"
date: "2/13/2021"
output: html_document
---

---

# Instructions
Create a word cloud using the `text` column from a real twitter dataset 
(sample.csv).

---

```{r setup, message = FALSE}
# Add your library below.
library(dplyr)
library(tidyr)
library(qdapRegex)
library(tm)
library(qdap)
library(wordcloud)
library(readr)


```

# Step 1. Load and Clean the Data
Load in the `sample.csv` dataset from this project's data folder. 
Please make sure you transform the document to lowercase, delete stopwords & numbers & punctuation (1 point).

```{r, "Step 1"}
# Write your code below.
sample <- read_csv("data/sample.csv")

# Getting the tweet text
twt_text <- sample$text

# Remove any URLs from the tweets
twt_txt_nourl <- rm_twitter_url(twt_text)

# Keeping only non-special characters
twt_txt_chrs <- gsub('[^A-Za-z]', " ", twt_txt_nourl)

# Converting to corpus
twt_corpus <- twt_txt_chrs %>%
                VectorSource() %>%
                Corpus()

# Converting test to lowercase
twt_corpus_lwr <- tm_map(twt_corpus, tolower)

# Removing Stop words
twt_corpus_stpwd <- tm_map(twt_corpus_lwr, removeWords, stopwords('english'))

# Removing additional spaces
twt_corpus_final <- tm_map(twt_corpus_stpwd, stripWhitespace)

```

---

# Step 2. Adjust the Stopwords
Add "can" and "just" to the stopwords list (1 point). 

```{r, "Step 2"}
# Write your code below.
# Creating a Vector of custom stop words
custom_stop <- c("can", "just")

# Removing custom stop words
twt_corpus_refined <- tm_map(twt_corpus_final, removeWords, custom_stop)

```

---

# Step 3. Adjust the Theme
Use five colors and "Accent" for color theme (1 point).

```{r, "Step 3"}
# Write your code below.
# Colorful word cloud
wordcloud(twt_corpus_refined, max.words = 100,
          colors = brewer.pal(5,"Accent"), scale = c(2.5,.5),
          random.order = FALSE)

```

---

# Step 4. Analysis
Does the word cloud convey the key points of the document? (1 point). 


> It shows that the tweets were about the airline Virgin America, and the 
  experience passengers had while traveling with this airline.

---

# Step 5. Compile
Submit the compiled file. 
The wordcloud may prevent you from compiling. If so, do the following: 

```
jpeg('p1.jpg')  
wordcloud(cloudFrame.t$word, cloudFrame.t$freq)  
dev.off()
```
This will make the graphs save in the current working directory. 

```{r, "Step 5"}
# Write your code below, if necessary.


```