---
title: "Week 11: Lab - Build Your Own Shiny App"
author: 'Ammar Ali'
date: '3/28/2021'
output: html_document
---

---

# Instructions
Create a scatter plot interactive application. In this application, use the 
census data from earlier in the book, and let the user determine the X-axis, 
the Y-axis, the color (optional), and the size based on what the user would 
like to see. To do this, you will need multiple-choice menus, one for each 
selection the user needs to make. You can do this by following the direction 
written in the text book. Enjoy. 

**For your information:**   
The data is available through this link: 

* http://www2.census.gov/programs-surveys/popest/tables/2010-2011/state/totals/nst-est2011-01.csv


Add all of your libraries that you use for this assignment here.
```{r setup}
# Add your library below.

library(shiny)
library(ggplot2)
```

---

# Step 1 - Build the UI
Build the UI below.

```{r, "Step 1"}
ui <- fluidPage(
  
  # Write your code here.
  selectInput("variable1", "X-axis:",
              c("july population" = "Pop2011",
                #"region of country" = "region",
                "change in population" = "popChange",
                "Percent change in population" = 
                'percentChange')),
  selectInput("variable2", "Y-axis:",
              c("july population" = "Pop2011",
                #"region of country" = "region",
                "change in population" = "popChange",
                "Percent change in population" = 
                'percentChange')),
  selectInput("variable3", "Size:",
              c("july population" = "Pop2011",
                #"region of country" = "region",
                "change in population" = "popChange",
                "Percent change in population" = 
                'percentChange')),
  plotOutput("plot")
  )
```

---

# Step 2 - Build the Server
Build the Server below.
```{r, "Step 2"}
server <- function(input, output) {
  readStates <- function(urlToRead = "http://www2.census.gov/programs-surveys/popest/tables/2010-2011/state/totals/nst-est2011-01.csv")
{
  #1. You need to read a URL, not a local file to your computer.
  
  #2. The file is a dataset on state populations (within the United States).
  #"http://www2.census.gov/programs-surveys/popest/tables/2010-2011/state/totals/nst-est2011-01.csv"
  
  states <- read.csv(url(urlToRead))
  
  #Step 2: Clean the dataframe: within Function 1
  
  #3. Note the issues that need to be fixed (removing columns, removing rows, changing column names).
  
  #4. Within your function, make sure there are 51 rows (one per state + the district of Columbia). Make sure there are only 5 columns with the columns having the following names (stateName, Census, Estimates, Pop2010, Pop2011).
  
  #5. Make sure the last four columns are numbers (i.e. not strings).

  ## Remove the first 8 rows
  states <- states[-1:-8, ]
  # Resets the rownames
  rownames(states) = NULL
  
  ## Move Puerto Rico one row up
  states[52, ] <- states[53, ] 
  
  ## Delete the last 5 rows
  states <- states[-53:-58, ]
  
  ## Remove the last 6 columns
  states <- states[ , -6:-10]
  
  # Fix the first column
  ## Change name to statenames
  names(states)[names(states) == "table.with.row.headers.in.column.A.and.column.headers.in.rows.3.through.4...leading.dots.indicate.sub.parts."] <- "StateName"
  
  ## Remove the period
  states$StateName <- gsub("\\.", "", states$StateName)
  
  # Fix the second column
  ## Change name to Census
  names(states)[names(states) == "X"] <- "Census"
  
  # Remove the comma and change to numeric
  states$Census <- as.numeric(gsub(",", "", states$Census))
  
  # Fix the third column
  names(states)[names(states) == "X.1"] <- "Estimates"
  states$Estimates <- as.numeric(gsub(",", "", states$Estimates))
  
  # Fix the fourth column
  names(states)[names(states) == "X.2"] <- "Pop2010"
  states$Pop2010 <- as.numeric(gsub(",", "", states$Pop2010))
  
  # Fix the fifth column
  names(states)[names(states) == "X.3"] <- "Pop2011"
  states$Pop2011 <- as.numeric(gsub(",", "", states$Pop2011))
  
  return(states)
  
}
  
 # Write your code here.
  dfStates <- readStates()
  dfStates <- dfStates[dfStates$stateName != "District of Columbia", ]
  dfStates <- dfStates[dfStates$stateName != "Puerto Rico", ]

  #dfStates$region <- state.region
  dfStates$stateName <- 
    tolower(dfStates$stateName)
  dfStates$popChange <- dfStates$Pop2011 - dfStates$Pop2010
  dfStates$percentChange <- dfStates$popChange/dfStates$Pop2010 * 100
  
  output$plot <- renderPlot(
    ggplot(dfStates) +
      geom_point(aes_string(x = input$variable1, y = input$variable2, size = input$variable3))
  )
}

```

---

# Step 3 - Create the Shiny App
Deploy the shiny app.
```{r, "Step 3"}

# Write your code here.
shinyApp(ui = ui, server = server)

```
