---
title: 'Week 9: Lab - Using aRules on the Titanic dataset '
author: 'Ammar Ali'
date: '3/14/2021'
output: html_document
---

---

# Instructions
Use the Titanic dataset to explore descriptive statistics, functions, and
association rules. Download the titanic dataset `titanic.raw.rdata` from one
of two locations:

- https://github.com/ethen8181/machine-learning/tree/master/association_rule/R
- https://sites.google.com/a/rdatamining.com/www/data/titanic.raw.rdata?attredirects=1
 
Note that it is not a cvs file, but rather an RData workspace. So, to load the 
data (assuming you saved it to the project's data folder), one would do:

``load("data/titanic.raw.rdata")``

You need to look at titanic.raw (the name of the R dataset)

``t <- titanic.raw``

Now that you have the datafile, do some descriptive statistics, getting some 
extra practice using R. Take the Quiz for Step 1 and 2. However your submission 
**MUST** include the process of you calculating the following values in Step 1 and 2.

---

```{r setup, message = FALSE}
# Add your library below.
#install.packages("arules")
#install.packages("arulesViz")
library(arules)
library(arulesViz)
library(dplyr)
```

# Step 0 - Load the data
Using the instructions above, load the dataset and save it as `t`.

```{r, "Step 0"}
# Write your code below.
load("data/titanic.raw.rdata")
t <- titanic.raw


```
---


# Step 1 - Descriptive stats (0.5 point for each answer)

1. Compute the percentage of people that survived.
2. Compute the percentage of people that were children.
3. Compute the percentage of people that were female.
4. Finally, compute the percentage of people that were in first class.

```{r, "Step 1"}
# Write your code below.
summary(t)

# 1. % Survivors
(711 / 2201) * 100
# 2. % Children
(109 / 2201) * 100
# 3. % Female
(470 / 2201) * 100
# 4. % 1st Class
(325 / 2201) * 100

```

---

# Step 2 - More descriptive stats (0.5 point for each answer)


1. What percentage of children survived? Your answer should be written such as # 13.75% of children survived
2. What percentage of female survived?
3. What percentage of first-class people survived?
4. What percentage of third-class people survived?

```{r, "Step 2"}
# Write your code below.
t_children <- t %>% filter(Age == "Child")
summary(t_children)
(57 / 109) * 100
# 1. 52.29% of children survived

t_female <- t %>% filter(Sex == "Female")
summary(t_female)
(344 / 470) * 100
# 2. 73.19% of females survived

t_fc <- t %>% filter(Class == "1st")
summary(t_fc)
(203 / 325) * 100
# 3. 62.46% of 1st Class survived

t_tc <- t %>% filter(Class == "3rd")
summary(t_tc)
(178 / 706) * 100
# 4. 25.21% of 3rd class survived


```

---

# Step 3 - Writing a function (0.5 point for each answer)

## Step 3.1 - Function 1
Write a function that returns a new dataframe of people that satisfies the 
specified criteria of sex, age, class and survived as parameters. I'm giving you 
the answer for this question:

```
myfunction1 <- function(a,b,c,d){
df1 <- t[t$Class == a,] # filter the data that satisfied the criteria that "Class" = a
df2 <- df1[df1$Sex == b,] # filter the data that satisfied the criteria that "Sex" = b
df3 <- df2[df2$Age == c,] # filter the data that satisfied the criteria that "Age" = c
df4 <- df3[df3$Survived == d,] # filter the data that satisfied the criteria that "Survived" = d
return(df4)}
# test the function with a sample data
myfunction1("1st","Female","Adult","No")
```

```{r, "Step 3.1"}
# Write your code below.
myfunction1 <- function(a,b,c,d){
df1 <- t[t$Class == a,] # filter the data that satisfied the criteria that "Class" = a
df2 <- df1[df1$Sex == b,] # filter the data that satisfied the criteria that "Sex" = b
df3 <- df2[df2$Age == c,] # filter the data that satisfied the criteria that "Age" = c
df4 <- df3[df3$Survived == d,] # filter the data that satisfied the criteria that "Survived" = d
return(df4)}
# test the function with a sample data
myfunction1("1st","Female","Adult","No")


```

## Step 3.2 - Function 2
Write a function, using the previous function, that calculates the percentage 
(who lives, who dies) for a specified (parameters) of class, gender and age 
considering the entire number of data. The function passes four arguments. 
Include the following code properly in your function by improvising names of objects.

```
p <- nrow(df)/nrow(t) # calculate the percentage
```

```{r, "Step 3.2"}
# Write your code below.
myfunction2<- function(a,b,c){
df1 <- t[t$Class == a,] # filter the data that satisfied the criteria that "Class" = a
df2 <- df1[df1$Sex == b,] # filter the data that satisfied the criteria that "Sex" = b
df3 <- df2[df2$Age == c,] # filter the data that satisfied the criteria that "Age" = c
p <- nrow(df3[df3$Survived == "Yes",])/nrow(t[t$Class ==a & t$Sex==b & t$Age ==c, ]) # calculate the percentage
p*100
return(p)}
# test the function with a sample data
myfunction2("1st","Female","Adult")


```

## Step 3.3 - Use the function (male)
Use the function to compare age and third-class male survival rates. 

```{r, "Step 3.3"}
# Write your code below.
myfunction2("3rd", "Male", "Child")
myfunction2("3rd", "Male", "Adult")


```
People in which category are more likely to survive? 

> Boys in 3rd class were more 11% more likely to survive than men in 3rd class


## Step 3.4 - Use the function (female)
Use the function to compare age and first-class female survival rates.

```{r, "Step 3.4"}
# Write your code below.
myfunction2("1st", "Female", "Child")
myfunction2("1st", "Female", "Adult")


```
People in which category are more likely to survive? 

> Girls in 1st class had a 100% at survival, compared to women who only had a 97% chance.

---

# Step 4 - Use aRules (0.5 point for each answer)

1. Use aRules to calculate some rules (clusters) for the titanic dataset.
2. Visualize the results.
3. Pick the three most interesting and useful rules. Explain these rules using natural language. **Answer this in the space provided below.**
4. How does this compare to the descriptive analysis we did on the same data set? 
Think critically. What was possible using one method that was not possible using 
the other method? **Answer this in the space provided below.**

```{r, "Step 4.1"}
# Write your code below.
m1 <- apriori(t)
m1
plot(m1, method = "grouped")

```

**Answer part 3 and 4 below.**

> The first rule that stands out {Class=Crew, Survived=No}, where you can see that many of the adult crew members most likely did not survive. The second, {Class=3rd, Sex=Male} also seemed interesting to me as it intersected with the {Survived=No}. This also reflects what was found in the descriptive analysis.The 2 rules {Survived=No} intersecting with {Age=Adult} and {Sex=Male} also is informative as it shows how many adult male passengers did not survive that shipwreck.
> This second method allows for the observation of patterns in the data that might not be easily found using just descriptive analysis. 


